\documentclass{article}
\usepackage{graphicx} % Required for inserting images
\usepackage{xcolor}
\begin{document}
\title{Homework 0}
\author{via Gradescope}

- Homework 0 gives you an opportunity to practice submitting a homework assignment through Gradescope. Note that you must assign a page to each problem you submit. Gradescope has great YouTube videos available on how to submit homework. Failure to submit homework or select pages correctly will result in zeroes.

- Handwritten homework is OK. You do not have to type up your work.

- Problems assigned from the textbook are from the $5th$ edition.

- No late homework accepted. Lateness due to technical issues will not be excused.

Please give a brief response to each of the following questions or instructions.

1. (a) Write your name and your NYU NetID (i.e. your NYU email).

{\color{blue} My name is Yilun Ma and my NetID is ym2932}

(b) What is your major or intended major?

{\color{blue} My major is Mathematics}

(c) What was your most recent math course before this one and when did you take it?

{\color{blue} I took Calculus III and Linear Algebra last semester}

2. (a) Why are you taking this course?

{\color{blue} To learn more math and prepare for the honor probability}

(b) What do you hope to learn or to gain from this course?

{\color{blue} How to prove statements and cultivate my ability in sets, logics, combinatorics, etc}

(c) How did the first week of our course go for you? (Hard or easy, fast or slow, etc.)

{\color{blue} I believe the first-week course is not too hard and in medium speed}

3. Write out the following statement. Include your name and signature where indicated. I, PRINT YOUR NAME, have carefully read the syllabus, as briefly reviewed in our first class meeting, and I understand the structure of the course as stated in the syllabus, including the homework and exam structure. If and when I have a question about the course structure, I will read the syllabus prior to asking the professor a question.

{\color{blue}I, Yilun Ma, have carefully read the syllabus, as briefly reviewed in our first class meeting, and I understand the structure of the course as stated in the syllabus, including the homework and exam structure. If and when I have a question about the course structure, I will read the syllabus prior to asking the professor a question.}

SIGN YOUR NAME:  Yilun Ma
\end{document}